Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-05-11 code/rmarkdown/article.Rmd
+ improvements to main text - typos and rewriting sections
Former-commit-id: f9b83a03b9c3a961f850aecf61292aa5ec2ec647

## Introduction
<!--  main text -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that protected areas are cost-effective, plans for protected areas (prioritizations) are often generated using decision support tools to identify near-optimal solutions. Additionally, because reserves cannot preserve the entire range for a given species, the plans for multiple reserves are often designed simultaneously to ensure that the network of protected areas can facilitate biological processes (eg. gene flow between populations, seasonal migration). For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions. However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is an important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490; but see @r489 for alternative methods]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can secure different genetic lineages and avoid the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform prioritizations because they require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]. Thus prioritizations may fail to capture species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have been investigating surrogates to generate prioritizations that adequately preserve genetic variation--but without actually needing to utilize genetic data to do this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sampled of environmental conditions across the species' ranges, they aimed to generate prioritizations that preserved more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Surrogates for neutral variation have been based on variables that predict the level of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively). Although these variables may correlate with genetic variation, they may not be effective surrogates if the level of correlation is not high enough or due to spatial auto-correlative effects. Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively). We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations. We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets. If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

## Methods
### Study system
We used species distribution and genomic data collected by the IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Although AFLPs are not ideal as genomic representatives (cf. single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists. Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

We chose this data set for several reasons. First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories. To ensure that our results were relevant for conservation planners, we required a similar sized data set. This data set contains genetic data for a diverse range of plant species. Second, we required genetic data for all species to be collected under a standardized sampling scheme, and this data set used a grid to collect samples. Third, we required broad-scale environmental variation across the study area to test the effectiveness of environmental variables. This data set covers a large geographic domain that spans across attitudinal gradients. Thus this data set was particularly well suited for our study.

This data set has previously been used to address a range of different questions. Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486]. However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning. Taberlet _et al._ [-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species. They found that the prioritization based on species distributions was substantially different to that based on genomic data. Their results suggest that prioritizations based on species distribution data alone will be unable to secure intra-specific genetic variation.

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We used the grid cells used to collect samples by the IntraBioDiv project as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of cost, we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

To describe the geographic location of each grid cell, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates. Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation.

To describe the environmental characteristics of each grid cell, we used broad-scale climatic variables. We obtained data for 19 bioclimatic variables across the extent of the study area [approx. 1 km resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation.

### Adaptive and neutral genetic data
To investigate the effectiveness of the surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were under selection. We used two outlier detection methods to achieve this [as recomended by @r495]. We omitted the species \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below). The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480]. The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral. To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.

First, we used multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456] to identify loci under selection. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. To avoid falsely classifying loci as adaptive, we grouped individuals into genetic lineages for each species (Figures S3--S28). We used the number of genetic lineages in each species previously identified by Alvarez _et al._ [-@r478]. Since \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} was previously identified as containing only one population, it was not compatible with \texttt{BayeScan}. For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations. We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].  Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

Second, we used PCAs to identify outlier loci using methods implemented in the \texttt{pcadapt R} package [version 3.0; @r487; @r496]. To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis. For each of the 26 species, we first imputed missing data to permit further analysis. Specifically, the value for a given individual at a given locus was replaced with the mean frequency of that locus among all conspecifics. Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci). We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491]. Outlier loci were detected using a suitable FDR ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species. For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices. We used non-metric multi-dimensional scaling [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation separately for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until this was achieved (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.

The previous analyses resulted in an adaptive (if detected) and neutral genetic space for each species. For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space. Planning units that were closer together in this space are occupied by individuals with similar neutral genetic loci. Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space. By spreading out conservation effort across a given genetic space for a given species, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].

@@

## Statistical analyses
To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data. For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (the PC values from the bioclimatc data) and geographic position (the z-scaled XY coordinates of the units' centroids). We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space. If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space. All dissimilarity matrices calculated using Euclidean distances.

We then fit maximum likelihood population-effects (MLPE) models [implemented in the \texttt{ResistanceGA R} package; @r488]. These models explicitly accommodate the structure of dissimilarity matrices using random effects [@r494]. For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation. Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using enviuronmental conditions and adaptive genetic variation. To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and Bonferroni corrections.

@@

We fit two full generalized linear models (GLM) with a logit links. The first model was fit to the proportion of adaptive genetic variation secured in a prioritization, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritization, a categorical variable indicating the species for which the prioritization was generated, and an interaction term between these predictor variables. The second model was fit to the proportion of neutral genetic variation secured in a prioritization, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritization. Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritizations were generated, and an interaction term. We subject these models to a backwards step-wise term deletion routine to assess term significance. To assess the performance of these surrogates, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively). We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].

To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations in real-world conservation planning scenarios--we investigated the effectiveness of surrogate-based targets using three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. However, most large-scale prioritizations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises. We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. Based on the previous analysis, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the variation was secured for each species. To accommodate variation in the high number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations. We then counted the proportion of species that were sufficiently represented in terms of their adaptive and neutral genetic variation (genetic variation secured $> `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)`$ %) for each prioritization.

>>>>

## Introduction
<!--  main text -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. However, the resources available for conservation are limited. Thus protected areas need to be sited in places that achieve conservation goals for minimal cost. To ensure that the plans for protected areas (prioritizations) are cost-effective, they are often generated using decision support tools to identify near-optimal solutions. These prioritizations often contain plans for a network of multiple protected areas so they can preserve biological processes (eg. connectivity between populations). For instance, \texttt{Marxan} [@r22] uses amount-based targets to identify protected area networks that preserve a suitable amount of habitat for each species (eg. 100 km $^2$ of land occupied by the species needs to be secured in a reserve) and weights to penalize fragmented solutions (ie. the \textit{BLM} parameter). However, conservation planning exercises typically assume that all individuals in the same species are equivalent.

Intra-specific genetic variation is important for long-term species persistence. As a consequence, there has been increasing interest in designing prioritizations that preserve and cultivate this variation [@r67; @r3; @r61]. Although the strength of natural selection is a continuous force; broadly speaking, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that (significantly) affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490; but see @r489 for alternative methods]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3]. In contrast, neutral genetic variation is associated with loci that do not (significantly) affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for developments on using genetic data in conservation planning]--especially since most large-scale exercises involve hundreds of species. Thus prioritizations may fail to secure a sufficient proportion of the intra-specific genetic variation of  species, and ultimately fail to deliver long-term conservation outcomes.

Recently, conservation scientists have been investigating the use of surrogate data to generate prioritizations that adequately preserve intra-specific genetic variation--but without actually needing to utilize genetic data to acheive this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been used to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, these researchers aimed to generate prioritizations that would preserve more adaptive variation. However, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Surrogates for neutral variation have been based on variables that predict the level of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritizations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigate the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. First, we aim to verify if environmental and geographic variables correlate with broad-scale patterns in adaptive and neutral variation (respectively). Although these variables may correlate with genetic variation, they may not be effective surrogates if they have poor predictive ability or if these correlations are driven by spatial auto-correlation. Second, we aim to verify that environmental and geographic variables are effective surrogates for adaptive and neutral variation (respectively). We expect that prioritizations generated using surrogate-based targets will secure more genetic variation than prioritizations generated using only amount-based targets with the same number of planning units. Third, we aim to determine if the addition of surrogate-based targets to conservation planning exercises can result in more effective prioritizations. We expect that prioritizations generated using amount- and surrogate-based targets will sufficiently preserve the genetic variation of more species than those generated using only amount-based targets. Furthermore, we aim to test this hypothesis under several of different scenarios to see whether the additional constraints of real-world planning exercises exacerbate or nullify this effect. If these surrogates can improve the intra-specific genetic variation captured in a prioritization, then conservation planners may use these surrogates when genetic data is not available, and in turn deliver more effective protected areas.

## Methods
### Study system
We used species distribution and genomic data collected by the international IntraBioDiv project in the European Alps [@r451; see @r452 and @r478 for further explanation of data collection methods]. This data set contains information for 27 alpine plant species (Figure 1a). It was collected using a 20' longitude by 21' latitude grid (approx. 22.3 km $\times$ 25 km; Figure S1). Project members visited every second grid cell, and if a species was detected in a cell, samples were collected from three individuals. They genotyped samples using amplified fragment length polymorphisms [AFLP; @r453], and constructed matrices denoting the presence/absence of polymorphisms at loci for each species (mean `r round(mean(sapply(spp.StructureData.LST, structurer:::n.loci)),2)` $\pm$ `r round(sd(sapply(spp.StructureData.LST, structurer:::n.loci)),1)` SD markers genotyped per species; see Table 1 in @r478 for the number markers for each species). Thus the data set contains information describing the genomic properties of individuals for each species present in every second grid cell.

We chose this data set for several reasons. First, conservation planning exercises typically consider tens to hundreds of species with different evolutionary histories and life histories. To ensure that our results were relevant for conservation planners, we required a data set with a diverse set of species and a comparable number of species. The IntraBioDiv data set contains genetic data for 27 species with different evolutionary histories and life-history strategies. Second, we required genetic data for all species to be collected using a standardized sampling scheme, and this data set used a grid to collect samples. Third, we required the study area to contain broad-scale environmental variation to test the effectiveness of environmental variables. This data set spans across large attitudinal gradients. Although this data set contains AFLPs  which not ideal as genomic representatives (cf. single nucleotide polymorphisms; SNPs), no multi-species genomic data set with similar properties currently exists. Thus this data set was well suited for our study.

This data set has previously been used to address a range of different questions. Many studies have used this data set to explore patterns of adaptive [@r484; @r451; @r475; @r455; @r485] and neutral genetic variation [@r478; @r451; @r486]. However, to our knowledge, only a single study has used this data set to explore the effectiveness of a surrogate in the context of conservation planning. Taberlet _et al._ [-@r486] investigated whether prioritizations generated using the spatial distribution of a comprehensive set species (> 300) could adequately secure the intra-specific genetic variation for 27 species. They found that the prioritization based on species distributions was substantially different to the prioritization based on genomic data. Their results suggest that prioritizations based on species distribution data alone will be unable to secure intra-specific genetic variation. However, whether or not environmental and geographic variables can augment prioritizations remains untested.

### Landscape data
We examined the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation (respectively). We utilized the sampling grid used by the IntraBioDiv project as planning units to develop prioritizations. We used only planning units that contained at least a single individual ($n = `r nrow(grid.DF)`$) to reduce computational burden. To examine the potential effects of variation in cost (see below), we calculated the total human population density inside each planning unit [obtained at 1 km resolution from the Global Rural-Urban Mapping Project; \texttt{GRUMP V1}; @r472] and used this to represent acquisition cost (Figure 1b).

To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic: $\texttt{ESRI:102031}$), calculated their centroids, and extracted their two-dimensional coordinates. Thus we constructed two-dimensional geographic space--wherein each planning unit occupies a single point--as a potential surrogate for neutral genetic variation. We expected planning units further apart from each other in this geographic space to contain individuals with increasingly different patterns in their neutral loci.

To describe the environmental characteristics of each planning unit, we used broad-scale climatic variation. We obtained data for 19 bioclimatic variables across the extent of the study area [approx. 1 km resolution; @r33]. We used this data set because it is commonly to map habitat suitability in conservation planning [@r493]. These layers and the planning units were projected into an equal-area coordinate system (Europe Lambert Conformal Conic: $\texttt{ESRI:102014}$). To reduce dimensionality, the 19 bioclimatic variables were clipped to the planning units and subjected to a principal components analysis (PCA; using \texttt{ArcMap} version 10.2.2; Table S1). The first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components (PCs) cumulatively explained `r round(pca.DF[[4]][surrogate.params.LST[[MODE]]$number.components],1)` % of the variation and were used to generate new layers. We calculated the average value for each PC layer in each planning unit, and used these values to characterize climatic variation among the planning units (Figure S2). Thus we constructed a three-dimensional environmental space--wherein each planning unit occupies a single point--as a potential surrogate for adaptive variation. We expect planning units further apart from each other in this environmental space to contain individuals with increasingly different patterns in their adaptive loci.

### Adaptive and neutral genetic data
To investigate the effectiveness of our supposed surrogates for adaptive and neutral genetic variation, we first needed to identify which of the sampled loci--if any--were adaptive. We used two outlier detection methods to achieve this [an individual- and a population-level method as recommended by @r495]. We omitted the species \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} from all subsequent analyses because it was not compatible with one of the outlier detection methods (see below). The basic premise underpinning such methods is that neutral loci are expected to exhibit a specific level of variation among populations, and loci that deviate from this expectation should be under selection [reviewed in @r480]. The advantage of these methods--in contrast with environmental association analyses [@r489]--is that they do not use environmental data to identify loci under selection, which would have biased our analysis. Loci identified as outliers in both analyses were treated as adaptive and the rest were treated as neutral. To avoid false-positives, we omitted loci from both outlier detection methods where the global frequency of the minor allele was $> 0.1$ and assumed they were neutral.

First, we used multinominal-Dirichlet models implemented in \texttt{BayeScan} [version 2.1; @r456] to identify loci under selection. We adopted a similar methodology to Bothwell _et al._ [-@r455] and applied it to each of the species in the data set. Following this methodology, for each species, we first grouped individuals into genetic lineages to avoid falsely classifying loci as adaptive. We used the number of genetic lineages identified in a previous analysis of the data set [@r478]. Since \textit{`r gsub('_', ' ' , missing.species, fixed=TRUE)`} was found to only contain a single lineage/population, it was not compatible with these models. For each species, we used \texttt{Structure} [version 2.3.4; using same parameters as @r478 per run with `r structure.params.LST[[MODE]]$numruns` runs per species] and \texttt{ClumPP} [version 1.1.2; @r459] to group individuals into populations. We then ran $\texttt{Bayescan}$ for each species using these groupings [1:1 prior odds; `r bayescan.params.LST[[MODE]]$nbp` pilot runs; `r bayescan.params.LST[[MODE]]$burn` burn-in iterations; `r bayescan.params.LST[[MODE]]$n` post-burn-in iterations thinned by `r bayescan.params.LST[[MODE]]$thin` iterations; @r456].  Additionally, we omitted individuals from this analysis when their population membership was uncertain (maximum membership probability $< `r structure.params.LST[[MODE]]$probability.threshold`$). We ran `r bayescan.params.LST[[MODE]]$reps` replicates per species using a suitable false discovery rate [$q \leq `r bayescan.params.LST[[MODE]]$fdr`$; @r466].

Second, we used PCAs to identify outlier loci [using methods implemented in the \texttt{pcadapt R} package; version 3.0; @r487; @r496]. To enable comparisons between the two outlier detection analyses, we used the same individuals in this analysis as used in the \texttt{BayeScan} analysis. For each of the 26 species, we first imputed missing data to permit further analysis. Specifically, the missing value for a given locus for a given individual was replaced with the average frequency of that locus among all conspecifics. Next, we ran a PCA and extracted the minimum number of components needed to secure the over-arching population level variation among the loci (`r pcadapt.params.LST[[MODE]]$min.variance.explained*10` % of the variation in loci). We then used Mahalanobis distances [@r398] to compute _P_ values and then q-values [using the \texttt{qvalue R} package; @r491]. Outlier loci were detected using a suitable false discovery rate ($q < `r pcadapt.params.LST[[MODE]]$alpha.level`$).

After classifying loci as adaptive or neutral, we mapped the main gradients of the adaptive (if detected) and neutral genetic variation for each species. For each species, we discarded the population groupings, and partitioned adaptive and neutral loci into separate matrices. We applied separate non-metric multi-dimensional scaling analyses to each matrix [NMDS; implemented in the \texttt{vegan R} package; @r458] using Gower distances [@r464; using the \texttt{cluster R} package to accommodate sparsity; @r457] to derive continuous variables that described the main gradients of adaptive and neutral genetic variation (separately) for each species (Table S2). To ensure that the ordinations described a sufficient amount of the genomic variation, we ran successive scaling analyses with increasing dimensionality until a sufficient stress value was obtained (maximum stress value $\leq$ `r nmds.params.LST[[MODE]]$max.stress`; `r nmds.params.LST[[MODE]]$trymax` random starts for each analysis). Since each grid cell had up to three samples per species, we used the average of the ordinated values associated with the individuals in each grid cell to express the typical genomic characteristics of individuals in the cells.

The previous analysis resulted in an adaptive (if detected) and neutral genetic space for each species. For a given species, each planning unit occupied by the species was associated with a multi-dimensional point in the species' neutral genetic space. Planning units that were closer together in this space are occupied by individuals with similar polymorphisms in their neutral loci. Similarly, if the species was associated with adaptive loci, each planning unit occupied by the species was associated with a multi-dimensional point in the species' adaptive genetic space. By spreading out conservation effort across a given genetic space, and in turn securing planning units occupied by individuals with increasingly different polymorphisms, prioritizations can secure more genetic variation [see @r465 for discussion on attribute spaces].

@@

## Statistical analyses
To address the first aim--to determine if the environmental and geographic variables correlate with spatial patterns of adaptive and neutral genetic variation--we computed dissimilarity matrices between the planning units, and tested for correlations between matrices based on surrogate and genetic data. For each species, we computed two dissimilarity matrices between the planning units occupied by the species in terms of their environmental conditions (their coordinates in the environmental space) and geographic position (their coordinates in the geographic space). We then computed another dissimilarity matrix between the planning units using their values in the species' neutral genetic space. If the species was associated with adaptive loci, we also computed another dissimilarity matrix between the planning units using the species' adaptive genetic space. All dissimilarity matrices calculated using Euclidean distances.

We then fit maximum likelihood population-effects (MLPE) models [@r494; implemented in the \texttt{ResistanceGA R} package; @r488]. These models explicitly accommodate the structure of dissimilarity matrices using random effects. For each species, we fit a MLPE model to correlate the species' dissimilarity matrices based on geographic position and neutral genetic variation. Additionally, if the species was associated with adaptive loci, we also fit a MLPE model to correlate the species' dissimilarity matrices calculated using environmental conditions and adaptive genetic variation. All predictor variables were z-transformed prior to model fitting to improve convergence. To test if the surrogate variables explained a substantial amount of genetic variation, we conducted $\chi^2$ tests between each model and its corresponding null model, and applied Bonferroni corrections.

@@

We fit two full generalized linear models (GLM) with a logit links. The first model was fit to the proportion of adaptive genetic variation secured in a prioritization, and contained a continuous predictor variable measuring the proportion of environmental variation also secured in the prioritization, a categorical variable indicating the species for which the prioritization was generated, and an interaction term between these predictor variables. The second model was fit to the proportion of neutral genetic variation secured in a prioritization, and contained a continuous predictor variable assessing the proportion of geographic variation also secured in the prioritization. Similar to the previous model, this model also contained a predictor variable indicating the species for which the prioritizations were generated, and an interaction term. We subjected these models to a backwards step-wise term deletion routine to assess term significance. To gauge the performance of these surrogates for each species, we fit a suite of models correlating the proportion of adaptive and neutral genetic variation held in each prioritization and the environmental and geographic variation they also secured (respectively). We then computed the Cragg and Uhler's pseudo-R$^2$ value for each model [using the \texttt{pscl} R package; @r492].

To address the third aim--to determine if the use of surrogate-based targets can actually improve prioritizations--we investigated the effectiveness of surrogate-based targets under three scenarios with increasing levels of complexity. To assess the benefits of surrogate-based targets under the simplest of circumstances, our first scenario involved generating single-species prioritizations for each species. However, most large-scale prioritizations involve a comprehensive set of species. By forcing solutions to secure individuals in multiple communities, this may result in prioritizations that secure a greater proportion of intra-specific genetic variation for each species. To test for this effect, our second scenario involved generating multi-species prioritizations. Finally, most large-scale prioritizations also explicitly consider opportunity cost to ensure that prioritizations are cost-effective. Our third scenario involved generating multi-species prioritizations that satisfied targets for minimal opportunity cost.

For each scenario, we generated three types of prioritizations using (1) only amount-based targets, (2) both amount- and surrogate-based targets, and (3) both amount- and genetic-based targets. The purely amount-based prioritizations served as our baseline--representing the prioritizations typically generated in conservation planning exercises (eg. by \texttt{Marxan}). We used `r (rapr.params.LST[[MODE]]$scenario.analysis$amount.target*100)` % amount-based targets in all prioritizations to ensure an adequate proportion of habitat was secured for each species. Based on the results from the analyses used to address our second aim, we set surrogate-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100)` % and genetic-based targets as `r (rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100)` % to ensure that most of the genetic variation was secured for each species. To accommodate the fact that there were a large number of optimal solutions for the single-species amount-based prioritizations, we generated `r (rapr.params.LST[[MODE]]$scenario.analysis$single.species.amount.replicates)` replicate prioritizations. We computed the proportion of adaptive (if present) and neutral genetic variation secured for each species in the prioritizations.
